---
title: "Bayes.Rmd"
author: "Kristan D Ramirez"
date: "2/11/2020"
output: word_document
---

# Naïve Bayes 


## Conditional probability of the "Delay" given "Carrier = DL," "Day of Week=Saturday (7)," "Destination = LGA," and "Origin = DCA." 


```{r}

delaydata <- read.csv('/Users/kris/Documents/delay.csv')
#delaydata

#Label Status of flight:
delaydata$Status[delaydata$Status == '0'] <- 'On-Time'
delaydata$Status[delaydata$Status == '1'] <- 'Delayed'
#str(delaydata) # Verify.


# Label 'Day of Week':
delaydata$Day.of.Week[delaydata$Day.of.Week=='1'] <- 'Sun'
delaydata$Day.of.Week[delaydata$Day.of.Week=='2'] <- 'Mon'
delaydata$Day.of.Week[delaydata$Day.of.Week=='3'] <- 'Tue'
delaydata$Day.of.Week[delaydata$Day.of.Week=='4'] <- 'Wed'
delaydata$Day.of.Week[delaydata$Day.of.Week=='5'] <- 'Thu'
delaydata$Day.of.Week[delaydata$Day.of.Week=='6'] <- 'Fri'
delaydata$Day.of.Week[delaydata$Day.of.Week=='7'] <- 'Sat'

# Convert 'Day of Week' to a factor:
delaydata2 <- cbind(delaydata, Day.of.Week = as.factor(delaydata$Day.of.Week))
delaydata3 <-delaydata2[-1]
#str(delaydata3) # Verify.




# Convert 'Status' to a factor:
delaydata4 <- cbind(delaydata3, Status = as.factor(delaydata3$Status))
delaydata5 <- delaydata4[-4]
str(delaydata5) # Verify.




# Partition Data Sets into testing and training:
# 70% of data into trainging data set.
# 30% of data into testing data set.
library(caret)
set.seed(1234)
intrain <- createDataPartition(y=delaydata5$Status, p=0.7, list=FALSE) 
trainSet <- delaydata5[intrain, ]
testSet <- delaydata5[-intrain, ]

#summary(trainSet) 
#summary(testSet)
#str(trainSet) # 300 obs, 5 vars.
#str(testSet) # 128 obs, 5 vars.

library(e1071)
# Conditional Probability for calculation:
Naive <- naiveBayes(Status~., data=trainSet)
Naive 



# Prediction Function:
Predict <- predict(Naive, newdata = testSet, type='class')


# Accuracy of the  predicted model:
Actual <- testSet$Status


confusionMatrix(Actual, Predict)
# Accuray of the predicted model is 93.7%.
```


###### Calculate the conditional probability of the "Delay" given "Carrier = DL," "Day of Week=Saturday (7)," "Destination = LGA," and "Origin = DCA."

###### (Status = Delayed )  given (Carrier = DL, Day of Week = 7, Destination = LGA, Origin = DCA)


###### P(Carrier = DL)  --->   0.043

###### P(Day of Week = Saturday)  --->  0.086

###### P(Destination = LGA)  --->   0.478

###### P(Origin = DCA)   --->   0.478

###### P(Status = Delayed)  --->  0.076

## COMBINE :::: 0.043 X 0.086 X 0.478 X 0.478 X 0.076 = .0000642



###### P(Carrier = DL)  ---> 0.125

###### P(Day of Week = Saturday)  ---> 0.172

###### P(Destination = LGA)  ---> 0.424

###### P(Origin = DCA)  ---> 0.528

###### P(Status = On Time)  ---> 0.923


## COMBINE :::: 0.125 X 0.172 X 0.424 X 0.528 X 0.923 = .0044426


## P(Status = 1 ) |(Carrier = DL, Day of Week = 7, Destination = LGA, Origin = DCA) = (0.000134) / ( 0.000134 + 0.004012) = 0.032


## FINAL CALCULATION :: (.0000642) / (.0000642 + .0044426) =  0.01425 =)


#Calculate the conditional probability of the "Delay" given "Carrier = DL," "Day of Week=Saturday (7)," "Destination = LGA," and "Origin = DCA."
```{r}
## predict probabilities
pred.prob <- predict(Naive, newdata = testSet, type = "raw")
## predict class membership
pred.class <- predict(Naive, newdata = testSet)

#Calculate the conditional probability of the "Delay" given "Carrier = DL," "Day of Week=Saturday (7)," "Destination = LGA," and "Origin = DCA."
df <- data.frame(actual = testSet$Status, predicted = pred.class, pred.prob)
df[testSet$Carrier == "DL" & testSet$Day.of.Week == "Sat" & testSet$Destination == "LGA" & 
     testSet$Origin == "DCA",]



```





## DELAYS ON ORIGINAL DATA:
###### Delays in the original data show 15 occur on Monday, 2 delayed flights on Saturday, 14 on Sunday, and 1 on Thursday; about 7.4% of flights were delayed. 

```{r echo=TRUE, fig.ratina=300, fig.cap="Bar Plot showing delayed flights for original dataset.", fig.align='center'}
table(delaydata5$Day.of.Week, delaydata5$Status)
count <-table(delaydata5$Status=="On-Time", delaydata5$Day.of.Week)
# total of 428 flights, with 32 delays, and 396 on time. 

# Barplot of delayed flights for each day of the week:
barplot(count, main = "Weekday Flight Delays", xlab= "Day Of The Week", ylab= "Counts", col= 'white')
```

## ACTUAL DELAY ON TEST DATA: 

###### Monday show 5 delayed flights, Saturday: 1, and Sunday: 3. 9/127 flights are shown as delayed; 7.08%.

```{r echo=TRUE, fig.ratina=300, fig.cap="Bar Plot showing Actual delayed flights from test data set.", fig.align='center'}
table(testSet$Day.of.Week, Actual)
Act_count <- table(Actual=="On-Time", testSet$Day.of.Week)
#Act_count 
barplot(Act_count, main = "Weekday Flight Delays", xlab="Day of The Week", ylab="Counts", col= 'white')
```



## PREDICTED DELAYS: 

###### The prediction model shows one delay for Monday, while the reamining flights are predicted to be on-time. Out of 127 flights, .7% of flights (1/127) are predicted to be delayed. PREDICTED and ACTUAL delays are comparible; .07 vs 7.08%.  


```{r echo=TRUE, fig.ratina=300, fig.cap="Bar Plot showing the Predicted delayed flights on the test dataset.", fig.align='center'}
table(testSet$Day.of.Week, Predict)
Pred_count <- table(Predict=="On-Time", testSet$Day.of.Week)
#Pred_count # Monday has one delay. 
barplot(Pred_count, main = "Weekday Flight Delays", xlab="Day of The Week", ylab="Counts", col= 'white')


```