---
title: "Multivariate1.rmd"
author: "Kristan Ramirez"
date: "10/18/2019"
output:
  word_document: default
  html_document: default
---

Problem 1. 
a). How many rows are in n X 3 matrix? How many columns? Choose a number for n
then give an example in R. 

```{r}
# a). If n=6, then the matrix would = 6 X 3, which would = 6 rows and 3 columns
A <- matrix(c(11,12,-14,77,3,-2,14,12,8,1,-6,-33,23,-14,36,13,17,-7), ncol=3, byrow=T)
A
dim(A)
```

Problem 1.
b) Give an example of a 3 X 2 matrix in R. 

```{r}
# b) 3 X 2 would equal 3 rows, 2 columns 
B <-matrix(c(3,4,2,8,5,3), ncol=2, byrow=T)
B
dim(B)
```

Problem 2.
a)Is it solvable? Yes. Since the first set's number of columns matches the number of rows in the second set, it is solvable.

```{r}
Problem_Two_a <- matrix(c(2,-1,1,2),ncol=2, byrow=T)
# Problem_Two_a
Problem_Two_aa <-matrix(c(1,1,2,2,0,1), ncol=3, byrow=T)
# Problem_Two_aa
Problem_Two_a %*% Problem_Two_aa
```

Problem 2. 
b)Is it solvable? Yes.  Since the first set's number of columns matches the number of rows in the second set, it is solvable.

```{r}
Problem_Two_b <- matrix(c(1,0,2,1,-2,1),ncol=2, byrow=T)
# Problem_Two_b
Problem_Two_bb <- matrix(c(1,1,2,2,0,1),ncol=3, byrow=T)
# Problem_Two_bb
Problem_Two_b %*% Problem_Two_bb
```

Problem 2. 
c) Is it solvable? No, by rule, the number of columns in the first set would need to equal the number
of rows in the second set. Since this example's first set has two columns and the second has one row, this is not solvable. 

```{r}
# As stated above, this function will not work. Error: non-conformable arguments. 
# Problem_Two_c <- matrix(c(1,0,2,1,-2,1), ncol=2, byrow=T)
# Problem_Two_c
# Problem_Two_cc <- matrix(c(1,-1,0), ncol=3)
# Problem_Two_cc
# Problem_Two_c %*% Problem_Two_cc
```

Problem 3.
a) Find the results of  the multiplication (A)(A^-1). Is it solvable? Why? Answer and explain.
Yes, this is a solvable problem, as A is a square matrix; 2 X 2, AND it's determinant is not equal to zero. 

```{r}
Problem_Three_a <- matrix(c(2,1,1,2), ncol=2, byrow=T)
# Problem_Three_a
Problem_Three_a_inv= solve(Problem_Three_a)

Problem_Three_a_inv
# Or we can round. 
round(Problem_Three_a %*% Problem_Three_a_inv, 10)
```

Problem 3.
b) Find the results of the multiplication (A)(A^-1). Is it solvable? Why? Answer and explain.
No, this function is not solvable, as we see that the determinant of the matrix is equal to zero, meaning it is not invertible and is singular. 

```{r}
Problem_Three_b <- matrix(c(2,0,1,0),ncol=2, byrow=T)
Problem_Three_b
#Problem_Three_b_inv = solve(Problem_Three_b)
#Problem_Three_b_inv
# Let's find out why this doesn't work.
round(det(Problem_Three_b), 10)
# We see that the determinant of our matrix is 0, which means it is not invertible
```

Problem 4. 
a) Report the covariance and correlation matrix among husband responses (only focus on variables:
h1, h2, h3, h4). Explain your findings. What variables are the most correlated variables? Why?
 
```{r}
love <- read.csv("http://tiny.cc/isqs6350_hus_wife")
# love
```
```{r}
# Correlation of husband variables 1-4. 
corr_h = cor(love[,1:4])
corr_h
```
```{r}
# Covariance of husband variables 1-4. 
cov_h = cov(love[,1:4])
cov_h
```

Problem 4. 
b) Report the covariance and correlation matrix among wife responses (only focus on variables: 
w1, w2, w3, w4). Explain your findings. What variables are the most correlated variables? Why?
 
```{r}
# Covariance of wife variables 1-4. 
cov_w =cov(love[, 5:8])
cov_w
```
```{r}
# Let's try the shorter operation to convert covariance of wife variable 1-4, to correlation.
cov2cor(cov_w)
```
```{r}
# Let's try another method to check the prior work.
# Correlation of wife variables 1-4.
corr_w = cor(love[,5:8])
corr_w
# Same results. 
```

Problem 4.  
c) Find the Euclidean distance matrix for the first 10 rows of data (which refers to the data of the
first 10 couples). Present your matrix in 2 decimal places. 
 
```{r}

# love
love.s = scale(love[1:10,])
d.s = dist(love.s[1:10, ])
# d.s
round((d.s), 2)

```

Problem 4.
d) Given your distance matrix in part c, among those 10 couples, which two couples are the most 
similar? For example, you can say couple 3 and 7 are the most similar (dissimilar) because their 
distance is the smallest (or largest). 
 
Couples 1 and 5 are most similar in this data set, as they have the lowest distance of the couples, while couples 1 and 10
seem to be most dissmilar with a distance of 6.06; the highest of the data set. 
 

Problem 5. 
Convert this covariance matrix into the correspoding correlation matrix. 
```{r}
Cov = matrix(c(3.8778, 2.8110, 3.1480, 3.5062,
               2.8110, 2.1210, 2.2669, 2.5690,
               3.1480, 2.2669, 2.6550, 2.8341,
               3.5062, 2.5690, 2.8341, 3.2352), byrow= T, nrow=4)
# Cov
```

```{r}
# Converting covariance matrix to the corresponsing correlation matrix. 
cov2cor(Cov)
```

 

 
 