

---
output:
  word_document: default
  html_document: default
---
Kristan D. Ramirez
Homework Assignment #3. 

PROBLEM 1.
a) Perform principal component analysis using correlation matrix. You dont need to 
do data cleaning or fixing the direction of variables in the data. 

We see that Comp. 1 covers most of the variance (35%)
```{r}
police <- read.csv("http://tiny.cc/isqs6350_police_app")
#police
# Let's get the correlation matrix to scale the data.
#cor(police)

#OR

#princomp finds the eigen vectors of the correlation matrix (after being switched from it's default of cov)
summary(princomp(police, cor = T)) # Reports the relative cumulative total variances. 


```
PROBLEM 1.
b) What percentage of total variance is covered by the first two principal components? The data above reports that
51% of the total variance is covered in principal component 1 plus principal component 2. This is displayed in
the row labeled Cumulative Porportion. The additional variance following, drops significantly. 



PROBLEM 1. 
c) Report the loading coefficients (eigenvector of the correlation matrix) of the first two principal components.
```{r}

police.pca <- princomp(police[, 1:2], cor = T) # Use cor to scale

summary(police.pca, loading = T)


```
PROBLEM 1. 
d) Can you attach a meaning to the first two principal components? Explain.
We were able to reduce the data to two variables/PCs; react and height. 
The data show the Comp. 1 column containing all positive values, thus can say it represents the overall scoring
for the categories React and Height. Loadings show that Comp1 and Comp2 are similar or highly related, meaning
that reaction time to visual stimuli is highly correlated to the police academy applicant's height. 



PROBLEM 2.
a) Perform the principal components using the correlation matrix. You don't need to do data cleaning or fixing the direction of variables in this data.

```{r}
crime <- read.csv("http://tiny.cc/isqs6350_crime", row.names= "STATE")
#head(crime)
crime_pc <- summary(princomp(crime, cor = T)) #set from default to correlation.
crime_pc

```

PROBLEM 2.
b) What percentage of total variance is covered by the first two principal components?
Component 1 covers 59% of the total variance, while Component 2 covers 18%, making up a combined 76%
which is easily identified by reviewing row 'Cumulative Proportions', or combining the totals for row 'Proportion
of Variance'.  

PROBLEM 2. 
c) Contruct the biplot graph of the crime data. Interpret the resulting biplot graph
for "MISSISSIPPI", "NEVADA", and "HAWAII". You can validate your conclusions by looking at the actual standardized (scaled) data values. 

If we draw a perpendicular line from the 3 states to the variable 'Murder', we see that Nevada has a 
higher rate of that particular crime, than the other two, but Hawaii has a notably significantly lower rate than 
the other two; Mississippi and Nevada. 


```{r}
#crime
states <- match(c("MISSISSIPPI", "NEVADA", "HAWAII"), rownames(crime))

biplot(crime_pc$scores[states, 1:2], crime_pc$loadings[,1:2], cex = c(0.6,.4))

```
